---
title: "API HW"
subtitle: "Cynthia Budiyanto"
format:
  pdf: default
editor_options: 
  chunk_output_type: console
---

```{r}
#| include: FALSE

library(tidyverse)
library(stringr)
library(httr2)
library(httr)
library(tidycensus)
```

### On Your Own - OMDB

1. Build a data set for a collection of movies by completing the **FILL IN** sections in the code below:

```{r}
#| message: false
#| warning: false

# http://www.omdbapi.com/?t=ANATOMY+OF+A+FALL
# http://www.omdbapi.com/?t=Death+Bed%3A+The+Bed+That+Eats
# http://www.omdbapi.com/?t=1998
# http://www.omdbapi.com/?t=Extremely+Loud+%26+Incredibly+Close
# http://www.omdbapi.com/?t=Fast+%26+Furious+Presents%3A+Hobbs+%26+Shaw

movies <- c("http://www.omdbapi.com/?t=ANATOMY+OF+A+FALL&apikey=", "http://www.omdbapi.com/?t=Death+Bed%3A+The+Bed+That+Eats&apikey=", "http://www.omdbapi.com/?t=1998&apikey=", "http://www.omdbapi.com/?t=Extremely+Loud+%26+Incredibly+Close&apikey=", "http://www.omdbapi.com/?t=Fast+%26+Furious+Presents%3A+Hobbs+%26+Shaw&apikey=")

# Set up empty tibble
omdb <- tibble(Title = character(), Rated = character(), Genre = character(),
       Actors = character(), Metascore = character(), imdbRating = character(),
       BoxOffice = character())

# Use for loop to run through API request process 5 times,
#   each time filling the next row in the tibble
#  - can do max of 1000 GETs per day
for(i in 1:5) {
  url <- str_c(movies[i], "e39b58ad")
  Sys.sleep(0.5)
  onemovie <- GET(url)
  details <- content(onemovie, "parse")
  omdb[i,1] <- details$Title
  omdb[i,2] <- details$Rated # rating
  omdb[i,3] <- details$Genre # genres (single string - could be more than one)
  omdb[i,4] <- details$Actors  # actors (single string - could be more than one)
  omdb[i,5] <- details$Metascore  # metascore (be sure it is numeric)
  omdb[i,6] <- details$imdbRating  # imdb rating (be sure it is numeric)
  omdb[i,7] <- details$BoxOffice  # box office (be sure it is numeric)
}

library(janitor)

omdb |>
  clean_names() |>
  mutate(metascore = parse_number(metascore),
         imdb_rating = parse_number(imdb_rating),
         box_office = parse_number(box_office))

#  could use stringr functions to further organize this data - separate 
#    different genres, different actors, etc.  But don't need to now.
```

### Navigating API documentation using httr

Navigate to the [Census API user guide](https://www.census.gov/data/developers/guidance/api-user-guide.html) and click on the "Example API Queries" tab.

Let's look at the Population Estimates Example and the American Community Survey (ACS) Example. These examples walk us through the steps to incrementally build up a URL to obtain desired data. This URL is known as a web API **request**. 

https://api.census.gov/data/2019/acs/acs1?get=NAME,B02015_009E,B02015_009M&for=state:*

- `https://api.census.gov`: This is the **base URL**.
    - `http://`: The **scheme**, which tells your browser or program how to communicate with the web server. This will typically be either `http:` or `https:`.
    - `api.census.gov`: The **hostname**, which is a name that identifies the web server that will process the request.
- `data/2019/acs/acs1`: The **path**, which tells the web server how to get to the desired resource.
    - In the case of the Census API, this locates a desired dataset in a particular year.
    - Other APIs allow search functionality. (e.g., News organizations have article searches.) In these cases, the path locates the search function we would like to call.
- `?get=NAME,B02015_009E,B02015_009M&for=state:*`: The **query parameters**, which provide the parameters for the function you would like to call.
    - We can view this as a string of key-value pairs separated by `&`. That is, the general structure of this part is `key1=value1&key2=value2`.

key      value
----     ------
get      NAME,B02015_009E,B02015_009M
for      state:*

Typically, each of these URL components will be specified in the API documentation. Sometimes, the scheme, hostname, and path (`https://api.census.gov/data/2019/acs/acs1`) will be referred to as the **[endpoint](https://en.wikipedia.org/wiki/Web_API#Endpoints)** for the API call.


### On Your Own - Census

2. Adapt the code in `hennepin_tidycensus` to write a function called `MN_tract_data` to give the user choices about year, county, and variables to pull off.  Show that `MN_tract_data(year = 2021, county = "Hennepin", variables = c("B01003_001", "B19013_001"))` works as expected.  Make sure it also works for other years, counties, and variables (e.g. B25077_001 is median home price and B02001_002 is number of white residents).

```{r}
#| message: false
#| warning: false

MN_tract_data <- function(year, county, variables) {
  tidycensus::get_acs(
    year = year,
    state = "MN",
    geography = "tract",
    variables = variables,
    output = "wide",
    geometry = TRUE,
    county = county,
    show_call = TRUE
  )
}

MN_tract_data(year = 2021, county = "Hennepin", variables = c("B01003_001", "B19013_001"))
```


3. Use your function from (2) along with `map` and `list_rbind` to build a data set for Rice county for the years 2019-2021.  Use your scraped data to plot trends in income over time and population over time.

```{r}
#| message: false
#| warning: false

years <- 2019:2021

rice19_21 <- map(years, ~MN_tract_data(year = .x, county = "Rice", variables = c("B01003_001", "B19013_001"))) |>
  list_rbind() |>
  as_tibble() |>
  rename(population = B01003_001E,
         median_income = B19013_001E)

year <- tibble(c(year = map_dbl(1:13, ~2019),
               year = map_dbl(14:29, ~2020),
               year = map_dbl(30:45, ~2021))) |>
  rename("year" = `c(...)`)

ricedf<- rice19_21 |>
  select(population, median_income) |>
  bind_cols(year) |>
  mutate(year = as.character(year))

ggplot(ricedf, aes(x = year, y = population)) +
  geom_boxplot() +
  geom_jitter(width = 0.2, color = "darkblue", alpha = 0.5) +
  theme_minimal() +
  labs(
    title = "Population Trend in Rice County (2019–2021)",
    x = "Year",
    y = "Population"
  )

ggplot(ricedf, aes(x = year, y = median_income)) +
  geom_boxplot() +
  geom_jitter(width = 0.2, color = "darkblue", alpha = 0.5) +
  theme_minimal() +
  labs(
    title = "Income Trend in Rice County (2019–2021)",
    x = "Year",
    y = "Income"
  )
```