---
title: "Strings: In-class Exercises (Part 2)"
format: pdf
editor_options: 
  chunk_output_type: console
editor: 
  markdown: 
    wrap: 72
---

You can download this .qmd file from
[here](https://github.com/joeroith/264_spring_2026/blob/main/03_strings_part2.qmd).
Just hit the Download Raw File button.

This uses parts of R4DS Ch 14: Strings and Ch 15: Regular Expressions
(both the first and second editions).

## Manipulating strings

str functions to know for manipulating strings

-   str_length()
-   str_sub()
-   str_c()
-   str_to_lower()
-   str_to_upper()
-   str_to_title()
-   str_replace() *not in 02_strings examples*

```{r}
#| message: false
#| warning: false

library(tidyverse)

#spotify <- read_csv("Data/spotify.csv") 
spotify <- read_csv("https://joeroith.github.io/264_spring_2026/Data/spotify.csv")

spot_smaller <- spotify |>
  select(
    title, 
    artist, 
    album_release_date, 
    album_name, 
    subgenre, 
    playlist_name
  )

spot_smaller


spot_smaller <- spot_smaller[c(5, 32, 49, 52, 83, 175, 219, 231, 246, 265), ]
spot_smaller
```

## Important functions for identifying strings which match

str_view() : most useful for testing 
str_subset() : useful for printing matches to the console 
str_detect() : useful when working within a tibble

1.  Identify the input type and output type for each of these examples:

```{r}
#| eval: false

str_view(spot_smaller$subgenre, "pop")
# input = chr vector
# output = chr vector using ANSI escapes

typeof(str_view(spot_smaller$subgenre, "pop"))
# input = any R object; in this case it is the result of str_view = vector
# output = character string

class(str_view(spot_smaller$subgenre, "pop"))
# input = any R object; in this case it is the result of str_view = vector
# output = character string

str_view(spot_smaller$subgenre, "pop", match = NA)
# based on the help document, match = NA will show all elements-- regardless if it matches the pattern or not. 
# input = chr vector
# output = chr vector using ANSI escapes

str_view(spot_smaller$subgenre, "pop", html = TRUE)
# input = chr vector
# output = vector in HTML formatting/widget

str_subset(spot_smaller$subgenre, "pop")
# input = vector
# output = vector
# should both be character

str_detect(spot_smaller$subgenre, "pop")
# input = vector
# vector = logical vector
```

2.  Use str_detect to print the rows of the spot_smaller tibble
    containing songs that have "pop" in the subgenre. (i.e. make a new
    tibble with fewer rows)

```{r}
spot_smaller |>
  filter(str_detect(subgenre, "pop"))

spot_smaller |>
  filter(str_extract(subgenre, "pop"))

spot_smaller |>
  filter(str_length(subgenre, "pop"))


x <- c("\u6c49\u5b57", "\U0001f60a")
str_view(x)
str_width(x)
str_length(x)
```

3.  Find the mean song title length for songs with "pop" in the subgenre
    and songs without "pop" in the subgenre.

```{r}
spot_smaller

spot_smaller |>
  mutate(sub_pop = 
           ifelse(str_detect(subgenre, "pop"), 
                  "Genre with pop", 
                  "Genre without pop")) |>
  group_by(sub_pop) |>
  summarise(mean_title_length = mean(str_length(title)))
```

Producing a table like this would be great:

```{r}
#| eval: false

# A tibble: 2 × 2
  sub_pop mean_title_length
  <lgl>               <dbl>
1 FALSE                18.6
2 TRUE                 13.6
```

Producing a table like this would be SUPER great (hint: `ifelse()`):

```{r}
#| eval: false

# A tibble: 2 × 2
  sub_pop           mean_title_length
  <chr>                         <dbl>
1 Genre with pop                 13.6
2 Genre without pop              18.6
```

4.  In the bigspotify dataset, find the proportion of songs which
    contain "love" in the title (track_name) by playlist_genre.

```{r}
#| message: false
#| warning: false

bigspotify <- readr::read_csv('https://raw.githubusercontent.com/rfordatascience/tidytuesday/master/data/2020/2020-01-21/spotify_songs.csv')

bigspotify
```

```{r}
bigspotify |>
  filter(!is.na(track_name)) |>
  select(track_name, playlist_genre, title_love) |>
  group_by(playlist_genre) |>
  summarise(prop_love = mean(title_love == "TRUE"))
```

## Matching patterns with regular expressions

|          |                                          |
|----------|------------------------------------------|
| `^abc`   | string starts with exactly "abc"         |
| `abc$`   | string ends with exactly "abc"           |
| `.`      | any character (helpful as a placeholder) |
| `[^abc]` | anything EXCEPT a or b or c              |

```{r}
#| eval: false
# Guess the output!

str_view(spot_smaller$artist, "^K")

str_view(spot_smaller$album_release_date, "01$")
str_view(spot_smaller$album_release_date, "01")
str_view(spot_smaller$album_release_date, "[01]")

str_view(spot_smaller$title, "^.. ")

str_view(spot_smaller$artist, "[^A-Za-z ]")
```

5.  Given the corpus of common words in stringr::words, create regular
    expressions that find all words that:

-   Start with “y”.
-   End with “x”
-   Are exactly three letters long.
-   Have seven letters or more.
-   Start with a vowel.
-   End with ed, but not with eed.
-   Words where q is not followed by u. (are there any in `words`?)

```{r}
#| eval: false

str_view(stringr::words, "^y") 

str_view(stringr::words, "x$")

str_view(stringr::words, "^...$")

str_view(stringr::words, "^.......")
str_view(stringr::words, "^.{7}")

str_view(stringr::words, "^[aiueo]") 

str_view(stringr::words, "[^e]ed$")

str_view(stringr::words, "q[^u]")

fruit1 <- c("apple", "banana", "pear", "pineapple")
fruit

str_subset(fruit1, "apple")
str_extract(fruit1, "apple")

fruit2 <- c("pineapple loveapple", "banana", "pearapple", "pineapple and apple")
str_view(fruit2, "apple")
str_subset(fruit2, "apple")
str_extract(fruit2, "apple")
str_extract_all(fruit2, "apple")
str_count(fruit2, "apple")
str_length(fruit2)


str_detect(fruit, "a")
str_detect(fruit, "^a")
str_detect(fruit, "a$")
str_detect(fruit, "b")
str_detect(fruit, "[aeiou]")
str_detect("aecfg", letters)

x <- "abcd ABCD 12345 -!@#%."
str_view(x, "\\d")
str_view(x, "\\d+")
str_view(x, "d")



```

```{r}
#| eval: false

# Try using str_view() or str_subset()

# For example, to find words with "tion" at any point, I could use:
str_view(words, "tion")
str_subset(words, "tion")
```

## More useful regular expressions:

|       |                              |
|-------|------------------------------|
| `\\d` | any number                   |
| `\\s` | any space, tab, etc.         |
| `\\b` | any boundary: space, ., etc. |

```{r}
#| eval: false

str_view(spot_smaller$album_name, "\\d")
str_view(spot_smaller$album_name, "\\s")
str_view(spot_smaller$album_name, "\\b")
```

Here are the regular expression special characters that require an
escape character (a preceding `\`): `^` `$` `.` `?` `*` `|` `+` `(` `)`
`[` `{`

For any characters with special properties, use `\` to "escape" its
special meaning ... but `\` is itself a special character ... so we need
two `\\`! (e.g. `\\$`, `\\.`, etc.)

```{r}
#| eval: false

str_view(spot_smaller$title, "$")
str_view(spot_smaller$title, "\\$")
```

6.  In bigspotify, how many track_names include a `$`? Be sure you print the track_names you find and make sure the dollar sign is not just in a featured artist!
    

# NOTE, NOT ANSWER

bigspotify |>
  filter(str_detect(track_name, "\\$")) |>
  filter(str_detect(track_name, [^feat])) |> will not work, because we are telling R that WE DO NOT WANT F, E, A, and T.
  filter(str_detect(track_name, "(feat|with).*\\$")) i want anything with "feat" or "with" THEN any number or characters afterward, but IT HAS TO HAVE A DOLLAR SIGN AFTER SEEING FEAT OR WITH.
  filter(str_detect())
  select(track_name, track_artist) |>
  drop_na(track_name) |>
  count(str_detect(track_name, "$")) |>


```{r}
# REAL ANSWER
bigspotify |>
  filter(str_detect(track_name, "\\$")) |>
  select(track_name, track_artist) |>
  print(n = Inf)

# how to do it with just for the title?

# if we only want the dollar sign on the featured artist.
bigspotify |>
  filter(str_detect(track_name, "(feat|with).*\\$")) |>
  select(track_name, track_artist) |>
  print(n = Inf)
```

7.  In bigspotify, how many track_names include a dollar amount (a `$` followed by a number).

```{r}
bigspotify |>
  filter(str_detect(track_name, "\\$\\d")) |>
  select(track_name, track_artist)
```


## Repetition

|         |                       |
|---------|-----------------------|
| `?`     | 0 or 1 times          |
| `+`     | 1 or more             |
| `*`     | 0 or more             |
| `{n}`   | exactly n times       |
| `{n,}`  | n or more times       |
| `{,m}`  | at most m times       |
| `{n,m}` | between n and m times |

```{r}
#| eval: false
str_view(spot_smaller$album_name, "[A-Z]{2,}")
str_view(spot_smaller$album_release_date, "\\d{4}-\\d{2}")
str_view(spot_smaller$album_release_date, "\\d{4}-(\\d{2})-\\1") #interesting

letters

str_length(letters)
str_length(NA)
str_length(factor("abc"))
str_length(c("i", "like", "programming", NA))
```

**Use at least 1 repetition symbol when solving 8-10 below**

8.  Modify the first regular expression above to also pick up "A.A" (in
    addition to "BEYONC" and "II"). That is, pick up strings where there
    might be a period between capital letters.

```{r}
#| eval: false
str_view(spot_smaller$album_name, "[A-Z]\\.?[A-Z]")
```


9.  Create some strings that satisfy these regular expressions and
    explain.

-   `"^.*$"`
-   `"\\{.+\\}"`

```{r}
#| eval: false
str_view("$uper", "^.*$")

str_view("", "^.*$")
```

Here are the regular expression special characters that require an
escape character (a preceding `\`): `^` `$` `.` `?` `*` `|` `+` `(` `)`
`[` `{`

```{r}
#| eval: false
str_view("{5}", "\\{.+\\}")

str_view("{this one work!}", "\\{.+\\}")

str_view("{this one does not work, because of the question mark!}", "\\{.?\\}")

str_view("this one does not", "\\{.+\\}")

str_view("hi", "\\{.+\\}")
```

double backslash... ESCAPE --> next character is the curly bracket. 
That means we are escaping the curly bracket. 
Why are we escaping it? Because the curly bracket is also a function to something else! We use curly bracket for repetition!
So here, we are asking it to find an actual curly bracket
Then the next thing is dot. Meaning, any THING! ANY CHARACTER!
Then, the plus sign --> meaning "1 or more". 


10. Create regular expressions to find all `stringr::words` that:

-   Start with three consonants.
-   Have two or more vowel-consonant pairs in a row.

```{r}
#| eval: false

str_view(stringr::words, "^[^aeiuo]{3}")

# has to be vowel then consonant pair
str_view(stringr::words, "([aiueo][^aiueo]|[aiueo][^aiueo]){2,}")

# either vowel-consonant or consonant-vowel pair.
str_view(stringr::words, "([aiueo][^aiueo]|[^aiueo][aiueo]){2,}")
```


## Useful functions for handling patterns

str_extract() : extract a string that matches a pattern 
str_count() : count how many times a pattern occurs within a string

```{r}
str_extract(spot_smaller$album_release_date, "\\d{4}-\\d{2}")

spot_smaller |>
  select(album_release_date) |>
  mutate(year_month = str_extract(album_release_date, "\\d{4}-\\d{2}"))

spot_smaller |>
  select(artist) |>
  mutate(n_vowels = str_count(artist, "[aeiou]"))
```

11. In the spot_smaller dataset, how many words are in each title?
    (hint: `\\b`)

```{r}
#WORDS not ALPHABET...

str_view(spot_smaller$title, "\\b[^ ]+\\b")

spot_smaller |>
  select(title) |>
  mutate(n_words = str_count(title, "\\b[^ ][A-Za-z]+(\\.[A-Za-z]+)*('[A-Za-z]+)*\\b"))

spot_smaller |>
  select(title) |>
  mutate(n_words = str_count(title, "\\b[^ \\d/]+\\b"))



str_view(spot_smaller$title, "\\b[^ \\d/]+\\b")
spot_smaller |>
  select(title) |>
  mutate(n_words = str_count(title, "\\b[^\\d/]+\\b"))


#initial code
spot_smaller |>
  select(title) |>
  mutate(n_words = str_count(title, "[^ ][A-Za-z]+"))

str_view(sentences, "^She|He|It|They\\b")

str_view(sentences, "^The\\b")
str_subset(sentences, "^The")
```

12. In the spot_smaller dataset, extract the first word from every
    title. Show how you would print out these words as a vector and how you would create a new column on the spot_smaller tibble. That is, produce this:

```{r}
# [1] "Hear"      "Run"       "Formation" "7/11"      "My"        "It's"     
# [7] "Poetic"    "A.D.H.D"   "Ya"        "Runnin"   
```

Then this:

```{r}
# A tibble: 10 × 2
#   title                                             first_word
#   <chr>                                             <chr>     
# 1 Hear Me Now                                       Hear      
# 2 Run the World (Girls)                             Run       
# 3 Formation                                         Formation 
# 4 7/11                                              7/11      
# 5 My Oh My (feat. DaBaby)                           My        
# 6 It's Automatic                                    It's      
# 7 Poetic Justice                                    Poetic    
# 8 A.D.H.D                                           A.D.H.D   
# 9 Ya Estuvo                                         Ya        
#10 Runnin (with A$AP Rocky, A$AP Ferg & Nicki Minaj) Runnin    
```

```{r}

str_extract(spot_smaller$title, "(^\\b[A-Z][a-z]+\\b|^\\b[A-Z].+\\b|^\\b\\d.+\\b)")




spot_smaller |>
  select(title) |>
  mutate(first_word = str_extract(title, "(^\\b[A-Z][a-z]+\\b|^\\b[A-Z].+\\b|^\\b\\d.+\\b)"))
```

13. Which decades are popular for playlist_names? Using the bigspotify
    dataset, try doing each of these steps one at a time!

1.   filter the bigspotify dataset to only include playlists that include something like "80's" or "00's" in their title.
2.  create a new column that extracts the decade
3.  use count to find how many playlists include each decade

4.  what if you include both "80's" and "80s"?
5.  how can you count "80's" and "80s" together in your final tibble?


```{r}
view(bigspotify)

#try
bigspotify |>
  select(playlist_name) |>
  filter(str_detect(playlist_name, "(\\d+['s]|\\d+'s)")) |>
  print(n = Inf)

bigspotify |>
  select(playlist_name) |>
  filter(str_detect(playlist_name, "(\\d+['s]|\\d+'s)")) |>
  mutate(decades = str_extract(playlist_name, "(\\d+'s|\\d+s|(\\d+s)-\\1)")) |>
  print(n = Inf)

bigspotify |>
  select(playlist_name) |>
  filter(str_detect(playlist_name, "(\\d+['s]|\\d+'s)")) |>
  mutate(decades = str_extract(playlist_name, "(\\d+'s|\\d+s|\\d+s-\\d+s)")) |>
  print(n = Inf)


#1
bigspotify |>
  select(playlist_name) |>
  filter(str_detect(playlist_name, "\\d+'s")) |>
  print(n = Inf)

#2
bigspotify |>
  select(playlist_name) |>
  filter(str_detect(playlist_name, "\\d+'s")) |>
  mutate(decades = str_extract(playlist_name, "\\d+'s"),
         decades = fct_recode(decades, "80's" = "808's",
                              "00's" = "2000's")) |>
  print(n = Inf)


#3: I don't actually understand what the sentence wants so...
bigspotify |>
  select(playlist_name) |>
  filter(str_detect(playlist_name, "\\d+'s")) |>
  mutate(decades = str_extract(playlist_name, "\\d+'s"),
         decades = fct_recode(decades, "80's" = "808's",
                              "00's" = "2000's")) |>
  count(decades, playlist_name) |>
  count(decades)

#4 & #5
bigspotify |>
  select(playlist_name) |>
  filter(str_detect(playlist_name, "(\\d+'s|\\d+s)")) |>
  mutate(decades = str_extract(playlist_name, "(\\d+'s|\\d+s)"),
         decades = fct_recode(decades, "80's" = "808's",
                              "00's" = "2000's",
                              "50's" = "1950s",
                              "10's" = "2010s"),
         decades = fct_collapse(decades, 
                                "80's" = c("80's", "80s"),
                                "90's" = c("90's", "90s"),
                                "70's" = c("70's", "70s")
                                )) |>
  count(decades, playlist_name) |>
  count(decades)

```

## Grouping and backreferences

```{r}
#| eval: false
# find all fruits with repeated pair of letters.  
fruit = stringr::fruit
fruit
str_view(fruit, "(..)\\1", match = TRUE)

# why does the code below add "pepper" and even "nectarine"?
str_view(fruit, "(..)(.*)\\1", match = TRUE)
# because of the (.*)! because that is saying that the (..) is followed by ANY CHARACTER 0 or more time... before repeating (..)
```

str_view(fruit, "(..)\\1", match = TRUE)
(..) any 2 characters
\\1 repeating what we have established above.
that's why we are getting "anan" "coco" "cucu" "juju" etc. 

str_view(fruit, "(..)(.*)\\1", match = TRUE)
(..) any 2 characters
(.*) any character 0 or more times. 
\\1 repeating our first position, which is the (..)

So, for example 
(.)(.)(..)\\2\\1
1. (.) any character
2. (.) any character again
3. (..) two character 
\\2 repeating THAT CHARACTER FROM POSITION 2
\\1 repeating THAT CHARACTER FROM POSITION 1


Tips with backreference:

-   You must use () around the the thing you want to reference.
-   To backreference multiple times, use `\\1` again.
-   The number refers to which spot you are referencing... e.g. `\\2`
    references the second set of ()

```{r}
#| eval: false
x1 <- c("abxyba", "abccba", "xyaayx", "abxyab", "abcabc")
str_view(x1, "(.)(.)(..)\\2\\1")
str_view(x1, "(.)(.)(..)\\1\\2")
str_view(x1, "(.)(.)(.)\\1\\2\\3")
```

14. Describe to your groupmates what these expressions will match, and
    provide a word or expression as an example:

-   `"(.)\\1\\1"`
-   `"(.)(.)(.).*\\3\\2\\1"`

For number 1, the expression will match one distinct character and repeat it twice (because we have two "\\1").
Example: ooo


For number 2, it will match a combination of three characters (as we have 3 "(.)", then the ".*" is saying that there will be any character 0 or more times then we are repeating that character from position 3, position 2, and position 1.
Example: abchellocba

Which words in `stringr::words` match each expression?

```{r}
#| eval: false
str_view(stringr::words, "(.)\\1\\1")
str_view(stringr::words, "(.)(.)(.).*\\3\\2\\1")
```

For the first one, none.
For the second one, "paragraph".

15. Construct a regular expression to match words in `stringr::words`
    that contain a repeated pair of letters (e.g. “church” contains “ch” repeated twice) but *not* match repeated pairs of numbers (e.g.
    507-786-3861).

```{r}
str_subset(stringr::words, "([a-z]{2}).*\\1")

str_view(words, "^(..).*\\1$")
```

16. Reformat the album_release_date variable in spot_smaller so that it
    is MM-DD-YYYY instead of YYYY-MM-DD. (Hint: str_replace().)

```{r}
spot_smaller |>
  mutate(album_release_date = str_replace(album_release_date,
      "(\\d{4})-(\\d{2})-(\\d{2})",
      "\\2-\\3-\\1"))
```

17. BEFORE RUNNING IT, explain to your partner(s) what the following R
    chunk will do:

```{r}
sentences |> 
  str_replace("([^ ]+) ([^ ]+) ([^ ]+)", "\\1 \\3 \\2") |> 
  head(5)

sentences |> 
  str_replace("([^ ]+) ([^ ]+) ([^ ]+)", "\\1 \\3 \\2") |> 
  head(5)

```

Will show the first five row. ([^ ]+) basically capturing word (or anything that is not a space). Then in the replacement argument, we swap the position of word 3 and 2.