---
title: "Strings: Extra Practice (Part 3)"
format: pdf
editor_options: 
  chunk_output_type: console
---
  
You can download this .qmd file from [here](https://github.com/joeroith/264_spring_2026/blob/main/04_strings_part3.qmd).  Just hit the Download Raw File button.

```{r}
#| message: false
#| warning: false

library(tidyverse)
library(rvest)
library(httr)
```


## On Your Own - Extra practice with strings and regular expressions

1. Describe the equivalents of ?, +, * in {m,n} form

? --> zero or one. {0,1}
+ --> one or more. {1,}
* --> zero or more. {0,}

2. Describe, in words, what the expression `"(.)(.)\\2\\1"` will match, and provide a word or expression as an example.


Gives us a combination of two characters (because we have two (.) then repeat the character from position 2, then repeat the character from position 1)


3. Produce an R string which the regular expression represented by `"\\..\\..\\.."` matches.  In other words, find a string `y` below that produces a TRUE in `str_detect`.

```{r}
y <- c(".w.o.w", "wrong", "w.r.ong.")
str_detect(y, "\\..\\..\\..")
```

4. Solve with `str_subset()`, using the words from `stringr::words`:

- Find all words that start or end with x.
- Find all words that start with a vowel and end with a consonant.
- Find all words that start and end with the same letter

```{r}
str_subset(stringr::words, "(^x|x$)")
str_subset(stringr::words, "^[aiueo].*[^aiueo]$")
str_subset(stringr::words, "^(.).*\\1$")
```

5. What words in `stringr::words` have the highest number of vowels? What words have the highest proportion of vowels? (Hint: what is the denominator?)  Figure this out using the tidyverse and piping, starting with `as_tibble(words) |>`.

```{r}
words

as_tibble(words) |>
  mutate(n_vowels = str_count(value, "[aiueo]"),
         n_alphabets = str_count(value, "[a-z]"),
         prop_vowel = n_vowels/n_alphabets) |>
  arrange(desc(prop_vowel))
  
```

Well, a has the "highest" proportion of vowels because it has one alphabet and that alphabet is a vowel :)

6. From the Harvard sentences data, use `str_extract` to produce a tibble with 3 columns:  the sentence, the first word in the sentence, and the first word ending in "ed" (NA if there isn't one).

```{r}
sentences

as_tibble(sentences) |>
  rename(Sentence = value) |>
  mutate(First = str_extract(Sentence, "[^ ]+"),
         ed = str_extract(Sentence, "\\b[A-Za-z]+ed\\b")) |>
  print(n = Inf)
```

7. Find and output all contractions (words with apostrophes) in the Harvard sentences, assuming no sentence has multiple contractions.

```{r}
as_tibble(sentences) |>
  rename(Sentence = value) |>
  filter(str_detect(Sentence, "\\b[A-Za-z]+'[A-Za-z].{0}\\b")) |>
  mutate(contractions = str_extract(Sentence, "\\b[A-Za-z]+'[A-Za-z].{0}\\b"))

#OR 

as_tibble(sentences) |>
  rename(Sentence = value) |>
  filter(str_detect(Sentence, "\\b[A-Za-z]+'[A-Za-z]\\b")) |>
  mutate(contractions = str_extract(Sentence, "\\b[A-Za-z]+'[A-Za-z]\\b"))       


#CODES         
str_extract(sentences, "\\b[A-Za-z]+'[A-Za-z].{0}\\b")
str_detect(sentences, "\\b[A-Za-z]+'[A-Za-z]\\b")
```


8. *Carefully* explain what the code below does, both line by line and in general terms.

```{r}
words
temp <- str_replace_all(words,                           # line 1
                        "^([A-Za-z])(.*)([a-z])$",       # line 2
                        "\\3\\2\\1")                     # line 3
as_tibble(words) |>                                      # line 4
  semi_join(as_tibble(temp)) |>                          # line 5
  print(n = Inf)                                         # line 6
```

In line one, we are creating a new dataset after using the str_replace_all function on the "words" string/vector. the str_replace_all function will replaces all matches, not only the first match.

In line 2, we are capturing any match that start with a letter (lower/upper case), then what comes after that is "." or basically anything..., then we capture the last letter (has to be lowercase).

Line 3 is changing the positon. last letter (has to be lowercase) will come first, then the middle techincally stays... then the beginning letter (upper/lowercase) will come last.

line 4 turns the words vector into a tibble (data frame.

then in line 5 the "as_tibble(temp)" part does the same thing as above, but then as we are using semi_join, we only keep only rows from words that exist in temp.

