---
title: "Knowledge Quiz 2"
subtitle: "Cynthia Budiyanto"
format:
  pdf: default
editor_options: 
  chunk_output_type: console
---

Guidelines: 

- No consulting with anyone else (other than clarification questions from Professor Roith)
- You may use only materials from this class (our class webpage, links on Moodle, our 3 online textbooks, files posted to the RStudio server)
- No online searches or use of large language models like ChatGPT

Pledge: 

I pledge my honor that on this quiz I have neither given nor received assistance not explicitly approved by the professor and that I am aware of no dishonest work.

- type your name here to acknowledge the pledge: Cynthia Budiyanto

```{r}
#| message: false
#| warning: false

library(tidyverse)
library(nycflights13)
```


1. Here's is a crazy list that tells you some stuff about data science.  Give code that will produce **exactly** the following outputs. (There are four parts to this question.)

```{r}
kq2_data <- list(
  DRY = c("Do not", "Repeat", "Yourself"),
  knowledge = c("Knowledge", "is", "Power"),
  nyc = flights,
  integers = 1:25,
  doubles = sqrt(1:25),
  toppings = list(sauce = c("marinara", "bbq", "creamy garlic"),
               cheese = c("mozzarella", "goat", "feta"),
               meat = c("pepperoni", "sausage", "chicken", "none"),
               veggie = c("peppers", "onions", "mushrooms", "pineapple"),
               hold = "no"),
  lostpts_to_perc = cbind(pts_lost = 0:50, 
                          perc = (50 - 0:50) / 50)
)

str(kq2_data)
```

a) 

[1] "Power"     "is"        "Knowledge"

```{r}
kq2_data[[2]][c(3,2,1)]
```

b) 
# A tibble: 6 × 3
# Groups:   dest [1]
  dest  carrier mean_arr_delay
  <chr> <chr>            <dbl>
1 MSP   9E                8.09
2 MSP   DL                4.04
3 MSP   EV               10.5 
4 MSP   MQ                9.56
5 MSP   OO               -2   
6 MSP   UA               -5.5 

```{r}
kq2_data[[3]] |>
  select(dest, carrier, arr_delay) |>
  filter(dest == "MSP") |>
  filter(!is.na(arr_delay)) |>
  group_by(dest, carrier) |>
  summarise(mean_arr_delay = mean(arr_delay))
```

c) Note: It must be one string and you must subset from kq2_data for the words "bbq", "goat", "chicken", and "mushroom" 

[1] "Give me a bbq pizza with goat cheese, chicken, and mushrooms."

```{r}
# Try 1

b <- kq2_data[[6]][[1]][2] #bbq

g <- kq2_data[[6]][[2]][2] #goat

c <- kq2_data[[6]][[3]][3] #chicken

m <- kq2_data[[6]][[4]][3] #mushrooms

str_glue(
  "[1] ",
  "\"",
  "Give me a {b} ",
  "pizza with {g} ",
  "cheese, {c}",
  ", and {m}",
  ".\""
)


#Try 2

mystrings <- c(f1 = "Give me a", f2 = "pizza with", f3 = "cheese,", f4 = ", and")

answer <- c(mystrings[1], kq2_data[[6]][[1]][2], mystrings[2], kq2_data[[6]][[2]][2], mystrings[3], kq2_data[[6]][[3]][3], mystrings[4], kq2_data[[6]][[4]][3])

```

d) 

[1] 10.0  0.8

```{r}
as.numeric(kq2_data[[7]][11,])
```


2. Write a function called `test_genre_popularity()` that produces a single ggplot. It should allow a user to input:

  1) a tibble
  2) two song genres
  3) and a plot option to create either a boxplot or faceted histograms
  
The title of your plot should indicate the two genres being compared. done

The subtitle should provide the *unique* test statistic and p-value for the specific genres being used, these values should change in the subtitle when you use different genres in the function input 

(Hint: the `t.test()` function will automatically calculate these numbers). 


Finally, your function should return an error when the user provides any plot option other than "boxplot" or "histogram". done


See the image KQ2_plot.png for examples of what the function should output.

```{r}
bigspotify <- readr::read_csv('https://raw.githubusercontent.com/rfordatascience/tidytuesday/master/data/2020/2020-01-21/spotify_songs.csv')

bigspotify



newbs <- bigspotify |>
  filter(playlist_genre %in% c("rap", "rock"))

stats <- t.test(track_popularity ~ playlist_genre, data = newbs)
  
str(stats)
stats[[1]]

```

```{r}
#| error: TRUE

test_genre_popularity <- function(df, genre1, genre2, plot) {
  twogenresdf <- df |>
    filter(playlist_genre %in% c( {{ genre1 }}, {{ genre2 }} )) # since doing t.test must have two levels of grouping factor
  stats <- t.test(track_popularity ~ playlist_genre, data = twogenresdf) # cant pipe t.test
  if (!(plot %in% c("boxplot", "histogram"))) {
  stop("Please use boxplot or histogram as the plotting option")
    }
  if (plot == "boxplot") {
  ggplot(twogenresdf, aes(x = playlist_genre, y = track_popularity)) +
      geom_boxplot() +
      labs(
      title = paste("Comparing", genre1, "and", genre2, "genres by popularity."),
      subtitle = paste("Two sample t-test for difference in mean popularity scores: T = ", stats[[1]], "and p-value = ", stats[[3]]))
  }
  else if (plot == "histogram") {
  ggplot(twogenresdf, aes(x = track_popularity)) +
      geom_histogram() +
      facet_wrap(~playlist_genre) +
      labs(
      title = paste("Comparing", genre1, "and", genre2, "genres by popularity."),
      subtitle = paste("Two sample t-test for difference in mean popularity scores: T = ", stats[[1]], "and p-value = ", stats[[3]]))
  }
}


test_genre_popularity(df = bigspotify, 
                      genre1 = "rap", 
                      genre2 = "rock", 
                      plot = "boxplot")

test_genre_popularity(df = bigspotify, 
                      genre1 = "rap", 
                      genre2 = "rock", 
                      plot = "density")
```


3. One of the harder concepts in statistics is the understanding of what the percentage in a confidence interval really means. Many introductory students believe that 95% confidence is a statement about the likelihood of the population parameter to be within the interval bounds. While not totally off-base, a better way to think about this is that if we take 1000 of different samples we expect around 950 of those samples to create 95% CIs that contain the true population parameter. Let's explore this.

Write a for loop that takes 10,000 samples of size 45 from a normal distribution with a true population mean of 25 and a true standard deviation of 4.  done

Calculate the lower bound and upper bound of a 95% confidence interval for the mean using each sample. done

Check to see what proportion of simulated samples actually contain 25. (Spoiler - it should be close to 0.95)

Here are a couple hints and reminders:

- `rnorm(n = 20, mean = 0, sd = 1)` will generate a random sample of size 20 from a Normal distribution with mean 0 and standard deviation 1.
- The equation for a 95% confidence interval is the sample mean plus/minus 1.96 times sample standard deviation divided by the square root of sample size.

$$\bar{x} \pm 1.96 \times \frac{s}{\sqrt{n}}$$
- The mean of a vector of logicals will give you a proportion.


```{r}

ci <- vector("logical", 10000)
for (i in 1:10000) {
  r_sample <- rnorm(n = 45, mean = 25, sd = 4)
  s_mean <- mean(r_sample)
  s_sd <- sd(r_sample)
  lower_ci <- s_mean - 1.96 * (s_sd / sqrt(45))
  upper_ci <- s_mean + 1.96 * (s_sd / sqrt(45))
  ci[i] <- (lower_ci <= 25) & (upper_ci >= 25)
}

mean(ci)
```



